Predator clusters

### 2023-03-26 reports/01_predators.qmd
Clean up ice and reports

effort_stars <- effort_sf %>% 
  filter(nmi < 1000) %>% 
  rasterize_counts(count_col = "nmi",
                   res = 2.5e4)
map_lim <- sf::st_bbox(c(xmin = -65, xmax = -55, ymin = -65, ymax = -58),
                       crs = "EPSG:4326") %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.2)
ant_basemap(map_lim) + 
  stars::geom_stars(aes(fill = nmi), data = effort_stars) +
  scale_fill_distiller(palette = "PuRd", 
                       direction = 1, 
                       na.value = NA) +
  scale_x_continuous(breaks = seq(-65, -55, by = 5)) +
  scale_y_continuous(breaks = seq(-65, -58, by = 2)) +
  coord_ant(map_lim) +
  facet_wrap(vars(year)) +
  expand_limits(fill = 0)

@@

map_lim <- sf::st_bbox(c(xmin = -65, xmax = -55, ymin = -66, ymax = -60),
                       crs = "EPSG:4326") %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.1)

@@

map_lim <- sf::st_bbox(soi_sf) %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.2)

>>>>

effort_stars <- effort_sf %>% 
  filter(nmi < 1000) %>% 
  rasterize_counts(count_col = "nmi",
                   res = 2.5e4)
map_lim <- st_bbox(c(xmin = -65, xmax = -55, ymin = -65, ymax = -58),
                       crs = "EPSG:4326") %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.2)
ant_basemap(map_lim) + 
  stars::geom_stars(aes(fill = nmi), data = effort_stars) +
  scale_fill_distiller(palette = "PuRd", 
                       direction = 1, 
                       na.value = NA) +
  scale_x_continuous(breaks = seq(-65, -55, by = 5)) +
  scale_y_continuous(breaks = seq(-65, -58, by = 2)) +
  coord_ant(map_lim) +
  facet_wrap(vars(year)) +
  expand_limits(fill = 0)

@@

map_lim <- st_bbox(c(xmin = -65, xmax = -55, ymin = -66, ymax = -60),
                       crs = "EPSG:4326") %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.1)

@@

map_lim <- st_bbox(soi_sf) %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.2)

### 2023-03-26 reports/02_stations.qmd
Clean up ice and reports

library(targets)
library(tidyverse)

@@

zoop_buffers_10k <- sf::st_buffer(zoop_sf, 5e3)
zoop_buffers_15k <- sf::st_buffer(zoop_sf, 10e3)
zoop_buffers_20k <- sf::st_buffer(zoop_sf, 20e3)

map_lim <- sf::st_bbox(zoop_sf) %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.2)

>>>>

library(sf)
library(targets)
library(tidyverse)

@@

zoop_buffers_10k <- st_buffer(zoop_sf, 5e3)
zoop_buffers_15k <- st_buffer(zoop_sf, 10e3)
zoop_buffers_20k <- st_buffer(zoop_sf, 20e3)

map_lim <- st_bbox(zoop_sf) %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.2)

### 2023-03-26 reports/03_combined.qmd
Clean up ice and reports

zoop_sf <- latlon_to_sf(zoop, 
                        coords = c("dec.longitude", "dec.latitude")) %>% 
  sf::st_transform(ant_proj()) %>% 
  rename(year = Year)
zoop_buffers_20k <- sf::st_buffer(zoop_sf, 20e3)

map_lim <- sf::st_bbox(zoop_sf) %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.2)

@@

>>>>

zoop_sf <- latlon_to_sf(zoop, 
                        coords = c("dec.longitude", "dec.latitude")) %>% 
  st_transform(ant_proj()) %>% 
  rename(year = Year)
zoop_buffers_20k <- st_buffer(zoop_sf, 20e3)

map_lim <- st_bbox(zoop_sf) %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.2)

@@

```{r}
#| label: ice-map
#| column: screen
#| out-width: 100%

ant_basemap(map_lim) +
  geom_sf(aes(color = ice_cat), 
          predators_stations,
          size = 0.5) +
  facet_wrap(~ year) +
  theme(legend.position = "bottom")

```

### 2023-03-26 reports/04_pred_clus.qmd
Clean up ice and reports

total_stations <- n_distinct(zoop_sf$amlr.station)
pred_frac <- predators_stations %>%
  group_by(amlr.station, species) %>% 
  summarize(count = sum(count), .groups = "drop") %>%
  group_by(species) %>%
  summarize(stations_present = n(),
            station_frac = stations_present / total_stations,
            .groups = "drop") %>%
  arrange(desc(station_frac))
cutoff_row <- sum(pred_frac$station_frac >= 0.05)

@@

species_wide <- pivot_wider(predators_abundant,
                            names_from = species,
                            values_from = count,
                            values_fill = 0) %>% 
  column_to_rownames("amlr.station")
cluster_indval <- labdsv::indval(species_wide, clust)
indval_mtx <- cluster_indval$indval
indval_fmt <- apply(indval_mtx, 2, function(x) {
  ord <- order(x, decreasing = TRUE)
  result <- ifelse(x >= 0.25, 
         str_glue("{rownames(indval_mtx)} ({round(x * 100)}%)"),
         NA) %>% 
    `[`(ord) %>% 
    na.omit() %>% 
    paste(collapse = "<br>")
  ifelse(result == "", "None", result)
})
as_tibble(t(indval_fmt)) %>% 
  knitr::kable(format = "html", escape = FALSE) %>% 
  kableExtra::kable_styling()

@@

map_lim <- sf::st_bbox(pred_clust_sf) %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.2)

@@

```{r}
cluster_fn <- function(x, k) {
  bray_dist <- function(x) {
    vegdist(x, method = "bray")
  }
  list(cluster = cutree(hclust(bray_dist(x), method = "ward.D2"), k = k))
}
set.seed(931)
sightings_gap <- suppressWarnings(
  clusGap(sightings_mtx, 
                   FUN = cluster_fn,
                   K.max = 10, 
                   B = 500,
                   verbose = FALSE)
)
sightings_gap_df <- as_tibble(sightings_gap$Tab) %>% 
  mutate(
    k = seq_along(logW),
    is_optimum = case_when(
      k == maxSE(gap, SE.sim, method = "Tibs2001SEmax") ~ "tibsSEmax",
      k == maxSE(gap, SE.sim, method = "firstSEmax") ~ "firstSEmax",
      TRUE ~ "none"
  ))
kfirstmax <- with(
  sightings_gap_df,
  maxSE(gap, SE.sim, method = "firstmax")
)
ktibs <- with(sightings_gap_df, k[is_optimum == "tibsSEmax"])
firstSEmax_thr <- with(sightings_gap_df, gap[kfirstmax] - SE.sim[kfirstmax])
tibs_thr <- with(sightings_gap_df, gap[ktibs + 1] - SE.sim[ktibs + 1])
ggplot(sightings_gap_df, aes(x = k, y = gap)) +
  geom_line() +
  geom_pointrange(aes(ymin = gap - SE.sim, ymax = gap + SE.sim,
                      color = is_optimum)) +
  geom_segment(x = 3, xend = 4, 
               y = firstSEmax_thr, yend = firstSEmax_thr,
               color = "red",
               linetype = "dashed") +
  geom_segment(x = ktibs, xend = ktibs + 1, 
               y = tibs_thr, yend = tibs_thr,
               color = "blue",
               linetype = "dashed") +
  scale_color_manual(values = c(tibsSEmax = "blue",
                                firstSEmax = "red",
                                none = "black")) +
  theme_classic() +
  theme(legend.position = "none")
```

>>>>

total_stations <- n_distinct(zoop_sf$amlr.station)
pred_frac <- predators_stations %>%
  as_tibble() %>% 
  group_by(amlr.station, species) %>% 
  summarize(count = sum(count), .groups = "drop") %>%
  group_by(species) %>%
  summarize(stations_present = n(),
            station_frac = stations_present / total_stations,
            .groups = "drop") %>%
  arrange(desc(station_frac))
cutoff_row <- sum(pred_frac$station_frac >= 0.05)

@@

species_wide <- predators_abundant %>% 
  pivot_wider(names_from = species,
              values_from = count,
              values_fill = 0) %>% 
  column_to_rownames("amlr.station")
cluster_indval <- labdsv::indval(species_wide, clust)
indval_mtx <- cluster_indval$indval
indval_fmt <- apply(indval_mtx, 2, function(x) {
  ord <- order(x, decreasing = TRUE)
  result <- ifelse(x >= 0.25, 
         str_glue("{rownames(indval_mtx)} ({round(x * 100)}%)"),
         NA) %>% 
    `[`(ord) %>% 
    na.omit() %>% 
    paste(collapse = "<br>")
  ifelse(result == "", "None", result)
})
as_tibble(t(indval_fmt)) %>% 
  knitr::kable(format = "html", escape = FALSE) %>% 
  kableExtra::kable_styling()

@@

map_lim <- st_bbox(pred_clust_sf) %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.2)

@@

```{r}
#| label: optimal-clusters

cluster_fn <- function(x, k) {
  bray_dist <- function(x) {
    vegdist(x, method = "bray")
  }
  list(cluster = cutree(hclust(bray_dist(x), method = "ward.D2"), k = k))
}
set.seed(931)
sightings_gap <- suppressWarnings(
  clusGap(sightings_mtx, 
                   FUN = cluster_fn,
                   K.max = 10, 
                   B = 500,
                   verbose = FALSE)
)
sightings_gap_df <- as_tibble(sightings_gap$Tab) %>% 
  mutate(
    k = seq_along(logW),
    is_optimum = case_when(
      k == maxSE(gap, SE.sim, method = "Tibs2001SEmax") ~ "tibsSEmax",
      k == maxSE(gap, SE.sim, method = "firstSEmax") ~ "firstSEmax",
      TRUE ~ "none"
  ))
kfirstmax <- with(
  sightings_gap_df,
  maxSE(gap, SE.sim, method = "firstmax")
)
ktibs <- with(sightings_gap_df, k[is_optimum == "tibsSEmax"])
firstSEmax_thr <- with(sightings_gap_df, gap[kfirstmax] - SE.sim[kfirstmax])
tibs_thr <- with(sightings_gap_df, gap[ktibs + 1] - SE.sim[ktibs + 1])
ggplot(sightings_gap_df, aes(x = k, y = gap)) +
  geom_line() +
  geom_pointrange(aes(ymin = gap - SE.sim, ymax = gap + SE.sim,
                      color = is_optimum)) +
  geom_segment(x = 3, xend = 4, 
               y = firstSEmax_thr, yend = firstSEmax_thr,
               color = "red",
               linetype = "dashed") +
  geom_segment(x = ktibs, xend = ktibs + 1, 
               y = tibs_thr, yend = tibs_thr,
               color = "blue",
               linetype = "dashed") +
  scale_color_manual(values = c(tibsSEmax = "blue",
                                firstSEmax = "red",
                                none = "black")) +
  theme_classic() +
  theme(legend.position = "none")
```
